### 2020-09-13 manuscript/paper_BRM.Rmd
change intro to focus more on generative models

author: 
  - name          : "Malte Lüken"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "malte_lueken@arcor.de"
  - name          : "Šimon Kucharský"
    affiliation   : "1"
  - name          : "Ingmar Visser"
    affiliation   : "1"

@@

authornote: |
  This is an author note.

@@

Gaze data has been frequently classified by probabilistic models. Compared to discrete (hard) classification, these models compute the probability of each event for every sample of the raw eye-tracking data. They have the advantage that they learn parameters from the data, can adapt to the task- and individual-specific gaze signals, and can be easily applied online [@Kasneci2014, p. 324].    
One class of probabilistic models that are used in eye movement classification are hidden Markov models (HMMs). HMMs estimate a sequence of hidden states (i.e., a discrete variable that cannot be directly observed) that evolves parallel to the gaze signal. Each gaze sample depends on its corresponding state. Each state depends on the previous but not on earlier states of the sequence [@Zucchini2016].  
In the context of eye movements, HMMs are suitable probabilistic models because the hidden states can be interpreted as eye movement events and gaze data are dependent time series (i.e., one gaze sample depends on the previous). Further, HMMs can be viewed as unsupervised models that can learn the hidden states and parameters of the emission process from the observed data alone, and as such do not in principle need training labeled data as it would be the case for supervised machine learning. On this basis, several classification algorithms using HMMs have been developed: One instance is described in @Salvucci2000 and combines the HMM with a fixed threshold approach (named "Identification by HMM" [I-HMM]). Samples are first labeled as fixations or saccades, depending on whether their velocity exceeds a threshold, and then reclassified by the HMM. @Pekkanen2017 developed an algorithm that filters the position of gaze samples through naive segmented linear regression (NSLR). The algorithm uses an HMM to parse the resulting segments into fixations, saccades, smooth pursuits, and PSOs based on their velocity and relative angle (named NSLR-HMM). Another version by @Mihali2017 uses a Bayesian HMM to separate microsaccades (short saccades during fixations) from motor noise based on sample velocity (named "Bayesian Microsaccade Detection" [BMD]). Moreover, @Houpt2018 developed a hierarchical approach that describes sample velocity and acceleration through an autoregression (AR) model, computes the regression weights through an HMM, and estimates the number of events with a beta-process (BP) from the data (named BP-AR-HMM).  

@@

HMMs have another property that has been rarely used in the context of eye movement classification: They can serve as a generative model for gaze data. To our knowledge, only @Mihali2017 used this property by building an explicit generative model for drift and microsaccades.  

>>>>

author: 
  - name          : "Malte Lüken"
    affiliation   : "1"
    corresponding : yes
    address       : "Speelmanstraat 7 1063ZC Amsterdam, The Netherlands"
    email         : "malte_lueken@arcor.de"
  - name          : "Šimon Kucharský"
    affiliation   : "1"
  - name          : "Ingmar Visser"
    affiliation   : "1"

@@

authornote: |
  We would like to thank Daan van Renswoude and Maartje Raijmakers for their helpful comments on earlier versions of this paper. Furthermore, we would like to acknowledge feedback by Karel Veldkamp and Phil Norberts in the early stage of this project.

@@

Most of these classification algorithms use non-parametric methods to differentiate between eye movement events. For instance, the classic "Velocity-Threshold" algorithm classifies samples with a velocity above a fixed threshold as saccades [@Stample1993]. More recently, supervised statistical learning techniques have been successfully used to parse eye movement data. For example, @Bellet2019 trained a convolutional neural network on eye-tracking data from humans and macaques and achieved classification results that were similarly accurate to those of human coders. However, these supervised methods have several disadvantages: First, they require labeled data (usually labeled by humans) as input. Second, they often rely on hyperparameters which are difficult to theoretically justify. Third, their parameters are usually difficult to interpret: Most supervised methods operate in a black-box fashion, which implies that they don't make any explicit assumptions about the data generating process. Oppositely, several algorithms using unsupervised techniques have been developed. They typically don't need to be trained on labeled data and are thus more flexible in classifying data from different individuals, tasks, or eye-trackers.  
Besides differentiating between supervised and unsupervised methods, algorithms can vary in whether they are explicitly modeling the data generating process and are thus able to simulate new data. To our knowledge, these *generative models* have been rarely used to classify eye movement data [cf. @Mihali2017, @Wadehn2019]. However, they come with various advantages and we will summarize the two most important for eye movement classification here [for an overview, see @Epstein2008].  
First, generative models allow to not only predict (classify) but also explain eye-tracking phenomena. For instance, they can inhere parameters that explain why oscillations only occur after but not before saccades. When the goal is to understand eye movements and improve their classification based on this understanding, this aspect is clearly an advantage over non-parametric or supervised methods. Second, generative models can challenge prevailing theoretical assumptions and bring up new research questions. For example, they might suggest that oscillations also occur before saccadic eye movements [as has been mentioned by @Nystrom2010] or that the assumption of discrete eye movements does not hold [as has been discussed in @Andersson2017; maybe others].  
One class of generative models that are used in eye movement classification are hidden Markov models (HMMs). HMMs estimate a sequence of hidden states (i.e., a discrete variable that cannot be directly observed) that evolves parallel to the gaze signal. Each gaze sample depends on its corresponding state. Each state depends on the previous but not on earlier states of the sequence [@Zucchini2016].  
In the context of eye movements, HMMs are suitable models because the hidden states can be interpreted as eye movement events and gaze data are dependent time series (i.e., one gaze sample depends on the previous). Further, HMMs can be viewed as unsupervised models that can learn the hidden states and parameters of the emission process from the observed data alone, and as such do not in principle need training labeled data as it would be the case for supervised methods.  
On this basis, several classification algorithms using HMMs have been developed: One instance is described in @Salvucci2000 and combines the HMM with a fixed threshold approach (named "Identification by HMM" [I-HMM]). Samples are first labeled as fixations or saccades, depending on whether their velocity exceeds a threshold, and then reclassified by the HMM. @Pekkanen2017 developed an algorithm that filters the position of gaze samples through naive segmented linear regression (NSLR). The algorithm uses an HMM to parse the resulting segments into fixations, saccades, smooth pursuits, and PSOs based on their velocity and relative angle (named NSLR-HMM). Another version by @Mihali2017 uses a Bayesian HMM to separate microsaccades (short saccades during fixations) from motor noise based on sample velocity (named "Bayesian Microsaccade Detection" [BMD]). Moreover, @Houpt2018 developed a hierarchical approach that describes sample velocity and acceleration through an autoregression (AR) model, computes the regression weights through an HMM, and estimates the number of events with a beta-process (BP) from the data (named BP-AR-HMM).  

@@
